---
title: "a4_task2_anna_talken"
author: "Anna Talken"
date: "11/14/2020"
output: html_document
---

```{r setup, include=FALSE, message = FALSE, warning = FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
library(here)
library(janitor)
library(lubridate)
library(ggbeeswarm)
library(kableExtra)
library(effsize)
library(broom)
```

## Introduction

This study is exploring data on juvenile snowshoe hares in Bonanza Creek from 1999 to 2012 as part of the US Long Term Ecological Research Network (Kielland, et. al 2017). Snowshoe hares are an important prey species in Alaska, but there is little quantitative information on their population dynamics. A few different analyses are performed here, including one to look at the change in the number of juvenile hares trapped annually from 1999 to 2012. Additionally, the relationship between juvenile hare weight and gender across different sites is explored, as well as the relationship between hind foot length and juvenile hare weight.

## Data Analysis

The data used in this report is derived from the Bonanza Creek LTER which is part of the US Long Term Ecological Research Network. In this report, descriptive statistics are analyzed, including mean, median, standard deviation and sample size count. Additionally, a two-sample t-test is used to determine any significant difference between weights of female vs. male juvenile hares. Finally, a regression model is used to look at the relationship between hind foot length and juvenile hare weight. 


## Exploratory Findings

```{r, message = FALSE, warning = FALSE, echo = FALSE}
snowshoe_hares <- read_csv(here("data", "bonanza_hares.csv")) #read in data
```


### 1. Annual juvenile hare trap counts

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap= "Figure 1: Annual number of juvenile hare trappings recorded in three different sites (Bonanza Mature, Bonanza Riparian, Bonanza Black Spruce) from 1999 to 2012"}
juvenile_hare_count <- snowshoe_hares %>% 
  filter(age == "j") %>% #filter to only include juvenile hares
  mutate(date = mdy(date)) %>% # Converted date to class "date"
  mutate(year_new = year(date)) %>% # Created a new year column
  count("j", year_new, sort = TRUE)  # Use the count function to count juvenile hares by year and SORT to order by number
 
# Create a ggplot to show the total number of hares per year
ggplot(data = juvenile_hare_count, aes(x = year_new, y= n)) +
  geom_col(fill = "cyan4") +
  labs(x = "Year", y = "Number of Juvenile Hares") +
  theme_minimal() +
  scale_x_continuous ("Year", breaks = seq(1999, 2012, 1)) # format x axis to have all the years listed
```
The minimum count of juvenile hare trappings is 2 hares in 2012 and the maximum count occurred in 1999 and was 126 hare traps (Table 1). The mean number of juvenile hares trapped is 31.5 and the median is 18.5. Moving forward, I would gather data based on 'effort' (number of days and number of traps used) so that I could standardize the data by juvenile hares trapped per unit effort. For example, I would look at the number of hares caught and the number of traps used, and find the proportion of hares caught per traps used. This would help to standardize the data across years and give a more accurate portrayal of annual trends in juvenile hare trappings. 


Table 1. Descriptive statistics (mean, median, minimum, maximum and sample size) of juvenile hare trappings from 1999 to 2012.
```{r, echo = FALSE, message = FALSE, warning = FALSE}
hare_summary <- juvenile_hare_count %>% 
  summarize(mean_number = mean(n),
            median_number = median(n),
            min_number = min(n),
            max_number = max(n),
            n = n())
#Create a summary table with min, max, mean and median

#Descriptive stats
hare_summary %>% 
  kable(col.names = c("Mean", "Median", "Minimum", "Maximum", "Sample Size")) %>% 
   kable_styling(bootstrap_options = "striped", full_width = FALSE)
```




### 2. Visualize juvenile hare weights

```{r, echo = FALSE, warning = FALSE, messages = FALSE, fig.cap = "Figure 2. Male and female juvenile hare weights recorded in each of the three sites: Bonanza Black Spruce, Bonanza Mature and Bonanza Riparian."}
#Creating a beeswarm box plot to display juvenile hare weight by sex and location.

juvenile_hares <- snowshoe_hares %>% 
  filter(age == "j") %>% 
  select(weight, sex, grid)  %>% #selecting to just see sex, site and weight of juvenile hares
  mutate(grid = case_when(
    grid == "bonbs" ~ "Bonanza Black Spruce",
    grid == "bonmat" ~ "Bonanza Mature",
    grid == "bonrip" ~ "Bonanza Riparian"))

ggplot(data = juvenile_hares, aes(x = sex, y = weight)) +
  geom_beeswarm(aes(color = sex)) +
  facet_wrap(~grid) +
scale_color_manual(values = c("goldenrod2","cyan4")) +
  geom_boxplot(fill = NA, width = 0.2, outlier.color = NA) +
  stat_summary(fun=mean, 
               geom="point", 
               shape=20, 
               size=4, 
               color="black", 
               fill="black") +
  theme_minimal() +
  labs(x = "Site", 
       y = "Weight (g)", 
       title = "Weight of female vs. male juvenile hares in three different trapping sites")
```
Figure 2 illustrates the average weight differences between male and female juvenile hares across three different sites (Bonanza Black Spruce, Bonanza Mature, Bonanza Riparian). Overall, males tend to have a higher average weight than female hares. Both female and male juvenile hares in Bonanza Black Spruce appear to have higher average weights than the other two sites.

### 3. Juvenile weight comparisons (male and female snowshoe hares)

```{r, echo = FALSE, message = FALSE, warnings = FALSE}
#Creating a table containing mean, SD, sample size for male and female juvenile hares to compare weight

weight_comparison <- juvenile_hares %>% 
  group_by(sex) %>% 
    summarize(
    mean_weight = round(mean(weight, na.rm = TRUE),2),
    sd_weight = round(sd(weight, na.rm = TRUE),2),
    n = n())
```

Table 2. Descriptive statistics (mean, standard deviation and sample size) for the weight of both male (m) and female (f) juvenile hares from 1999 to 2012. 

```{r, echo = FALSE, message = FALSE, warnings = FALSE}
weight_comparison %>% 
  kable(col.names = c("Sex", "Mean", "SD", "Sample Size")) %>% 
   kable_styling(bootstrap_options = "striped", full_width = FALSE)
```

```{r, echo = FALSE, message = FALSE, warnings = FALSE, include = FALSE}
#T-test

f_hares <- juvenile_hares %>% 
  filter(sex == "f") %>% 
  pull(weight)

m_hares <- juvenile_hares %>% 
  filter(sex == "m") %>% 
  pull(weight)

#Plot histograms & qq-plots to check assumptions
hist(f_hares)
qqnorm(f_hares)
hist(m_hares)
qqnorm(m_hares)

# Female qq plot is relatively linear, male qq plot slightly deviates from a linear trend. 

#Run the t-test:
hares_ttest <- t.test(f_hares, m_hares)

#Get the tidy model results to call outputs in-line: 
hares_ttest_tidy <- tidy(hares_ttest)

#Get the means & SDs for each: 
male_mean <- mean(m_hares)
female_mean <- mean(f_hares, na.rm = TRUE)
male_sd <- sd(m_hares)
female_sd <- sd(f_hares)

#Get the effect size: 
hares_effsize <- cohen.d(f_hares, m_hares, na.rm = TRUE)
```

When comparing weights (g) between male and female juvenile hares, males have a higher average weight (`r round(male_mean,2)` $\pm$ `r round(male_sd, 2)` than female hares (`r round(female_mean,2)` $\pm$ `r round(female_sd, 2)`. The difference in means is `r round(male_mean, 2) - round(female_mean, 2)` mm (a `r round((male_mean - female_mean)/((male_mean + female_mean)/2)*100, 2)`% difference), was shown to be significant (Welch's two-sample t-test: t(`r round(hares_ttest_tidy$parameter,2)`) = `r round(hares_ttest_tidy$statistic,2)`, p < 0.01), and the effect size is small (Cohen's *d* = `r round(hares_effsize$estimate,2)`).


### 4. Relationship between juvenile weight and hind foot length

Here I am exploring the relationship between juvenile hare weight (g) and hind foot length (mm). 

```{r, echo = FALSE, message = FALSE, warning = FALSE, fig.cap= "Figure 3. A comparison of the hind foot length and body weight of juvenile snowshoe hares."}
#Creating a new data frame that includes only the juvenile hare weight and hind foot length
hare_hind <- snowshoe_hares %>%  
  filter(age == "j") %>%  # only include juvenile hares
  select(weight, hindft) #only include hind foot and weight data

#Plot the comparison of hind foot vs. weight
ggplot(data = hare_hind, aes(x= weight, y = hindft)) +
  geom_point(color = "cyan4") +
  geom_smooth(method = "lm", se = FALSE, color = "gray30") +
  ggpubr:: stat_regline_equation( label.x = 1250, label.y = 80) + #add regression line equation
  labs(x = "Weight (g)",
       y = "Hind foot (mm)") +
  theme_minimal()
```

```{r, echo = FALSE, message = FALSE, warnings = FALSE, include = FALSE}
#Perform a linear regression
hare_hind_lm <- lm(hindft ~ weight, data = hare_hind)

#Get tidy versions of the model output
hare_hind_tidy <- tidy(hare_hind_lm)
hare_hind_glance <- glance(hare_hind_lm)

#See model diagnostics
plot(hare_hind_lm)

#Run a Pearson's r correlation
hare_hind_cor <- cor.test(hare_hind$weight, hare_hind$hindft)

#Tidy version of correlation output
hare_hind_cor_tidy <- tidy(hare_hind_cor)

```

The relationship between juvenile hare weight and hind foot length does not appear to be very linear. A regression model revealed that hind foot length and hare weight (g) do not have a strong linear relationship (p < 0.001, R^2^ = `r round(hare_hind_glance$r.squared,2)`) with an average slope of $\beta$ = `r round(hare_hind_tidy$estimate[2], 2)` g mm^-1^ (Figure 3). Juvenile hare weight and hind foot length are somewhat positively correlated, but not very strongly (Pearson's *r* = `r round(hare_hind_cor_tidy$estimate,2)`, p < 0.001). Diagnostic plots of the residual variance (not included) reveal heteroscedasticity. It is possible that combining the two sexes of the hares into one regression analysis influenced the variance of the residuals and weakened the linear relationship between hind foot length and weight. It would be useful to run two different regression analyses after separating the data by sex. 


## Summary

* There overall number of juvenile hare traps declines over the years from 1999 to 2012. 
* On average, juvenile male hares have a significantly greater weight than females, though the actual difference in weights is not that large (small effect size)
* The site "Bonanza Black Spruce" had juvenile hares with the greatest average weight for both female and male hares
* There is not a strong linear relationship between hind foot length (mm) and juvenile hare weight(g)

## Citations

Kielland, K., F.S. Chapin, R.W. Ruess, and Bonanza Creek LTER. 2017. Snowshoe hare physical data in Bonanza Creek Experimental Forest: 1999-Present ver 22. Environmental Data Initiative. https://doi.org/10.6073/pasta/03dce4856d79b91557d8e6ce2cbcdc14

